---
title: "Modeling Logistic Growth of Wild Fish Catch"
description: |
  Surprise, another post about fish! Here, I explore nonlinear least squares (NLS) as a method of predicting fish growth paramters.
author:
  - name: Dylan
    url: {}
date: 03-13-2021
output:
  distill::distill_article:
    self_contained: false
    code_folding: code
---
```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(janitor)
library(here)
library(boot)
library(gt)
library(patchwork)
library(broom)
library(nlstools)
```



##### Reading in Data

```{r, echo = FALSE}
fish_catch <- read_csv(here("_posts", "2021-03-13-post-2", "fish_catch.csv"), skip = 2) %>% 
  clean_names() %>% 
  drop_na() %>% 
  mutate(year = as.numeric(year), 
         wild_catch = as.numeric(wild_catch), 
         farmed_fish = as.numeric(farmed_fish), 
         total_fish_production = as.numeric(total_fish_production))
```


##### Exploratory Visualization

```{r, echo = FALSE}
three_part_plot <- ggplot(data = fish_catch) +
  geom_line(aes(x = year, y = wild_catch), color = "goldenrod") +
   geom_line(aes(x = year, y = farmed_fish), color = "darkblue") +
   geom_line(aes(x = year, y = total_fish_production), color = "red")

three_part_plot
```
Here we see:
- Wild caught fish looks similar to a logistic equation. 
- Farmed fish looks exponential, with an X- intercept around 1960
- Total fish looks somewhere between logisitc and linear, but the late arrival of farmed fish gives this line a lot of noise. 


Let's also check the log of wild catch
```{r, echo = FALSE}
wild_catch_plot <- ggplot(data = fish_catch)+
  geom_line(aes(x = year, y = log(wild_catch)), 
             color = "darkblue")

wild_catch_plot
```


##### Logistic Growth equation

**$P(t)=\frac{K}{1+Ae^{-rt}}$**, where

- $K$ is the carrying capacity
- $A$ is $\frac{K-P_0}{P_0}$
- $r$ is the growth rate constant

Here, our actual variable from the data is time (t.) This we won't change. The other parameters we will try and estimate!

To get an accurate time (t) measurement, let's set the year 1950 as 0. This mean 1965 with give a t = 20 value. 
```{r, echo = TRUE}
fish_time <- fish_catch %>% 
  mutate(time = year-1950)
```


##### Building a model with a year-based df 

```{r, echo = TRUE}
df_nls <- nls(wild_catch ~ K/( 1 + A*exp(-r*time)),
              data = fish_time,
              start = list(K = 180, A = 17, r =0.17),
              trace = TRUE)

 summary(df_nls)

p_predict <- predict(df_nls)
```

This model gives us:
- $K = 1.003 x 10^2$
- A = 4.316
- r = $6.989 x 10^-2$


```{r, echo=FALSE}
p_predict <- predict(df_nls)

fish_complete <- data.frame(fish_time, p_predict)

```


We get a final equation of: 

**$P(t)=\frac{100.278}{1+4.316e^{-0.06989t}}$**,

Alright, this is great information on the logistic growth equation. 

##### Final Visualization

I'll finish with a complete visualization of this curve, alongside our actual observations. 

```{r, echo = TRUE}
 final_plot <- ggplot(fish_complete) +
  geom_line(aes(x = year, y = p_predict)) +
  geom_point(aes(x = year, y = wild_catch),
             color = "cornflowerblue") +
  theme_bw() +
  labs(x = "Year", 
       y = "Wild Fish Catch (Millions of Tons)", 
       title = "Wild Fish Catch, Predicted v. Actual (1950-2012)") +
  scale_x_continuous(breaks = c(1950, 1960, 1970, 1980, 1990, 2000, 2010))
  
final_plot
```
*Figure 1: Wild Fish Catch, Predicted v. Actual (1950-2012): The logistic line, shown in black, marks the expected fish catch based on a logistic growth model that uses time (years,) carrying capacity, and intrinsic growth rate. The actual catch data from each year are shown in blue dots. All counts are in millions of tons*
