---
title: "Spatial Data Exploration"
description: |
  Here is my first exploration with raster data. I plan to explore the possibilities of Marine Spatial Planning moving forward, so why not start with whales and dolphins?
author:
  - name: Dylan
    url: {}
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning= FALSE)
library(tidyverse)
library(raster)
library(here)
library(rnaturalearth)
library(sf)
```

#### Whale (Cetation) Raster Work


Let's start by reading in the idfferetn .tif files and rasterizing them. We'll try using raster_stack, since there so many layers. 

```{r}
my_whale_files <- list.files(path = here("ca_cetaceans"), full.names = TRUE)


cetacean_raster <- raster::stack(my_whale_files)

# reclassified_stack <- reclassify(cetacean_raster, c(-Inf, 0.59,  0,  0.59,  Inf,  1))

cetacean_presence <- function(x, thresh = 0.6) {
y <- ifelse(x >= thresh, 1, NA) 
return(y)
}

## Calculating whale rasters with the function I made
whales_in_sea <- calc(cetacean_raster, fun = cetacean_presence)

## Plotting the rasterbrick + function
plot(whales_in_sea, col = 'lightblue')

## Converting rasterstack to dataframe, and raster_to_points

cetacean_df <- raster::rasterToPoints(whales_in_sea) %>%
  as.data.frame()
```

#### Making a *longer* df that lists all cetaceans by species, and them summarizing it. 

I use `pivot_longer` to have a species column with all 35 species, rather than 35 columns (one for each species.)

```{r}
cetacean_longer_df <- cetacean_df %>% 
  pivot_longer(cols =!  x:y,
               names_to = "species",
               values_to = "value") %>% 
  group_by(x,y) %>% 
  summarize(richness = sum(value, na.rm = TRUE))
```



#### Bringing in a background layer from the rnaturalearth package.
This includes a few different things:
- Bringing in world map data (which includes CA) from `rnaturalearth`
- Match the crs between this `rnaturalearth` and my `cetacean_raster` stack
- Map the earth layer to see it

```{r, results='hide'}
nat_earth_layer <- ne_countries(scale = "medium", returnclass = "sf")

## Checking the crs() of this layer
st_crs(nat_earth_layer)

## Making sure that the crs of my origional raster stack and the natural earth layer match up. 
nat_earth_layer <- st_transform(nat_earth_layer, st_crs(cetacean_raster))
 
st_crs(nat_earth_layer)


plot(nat_earth_layer)
```



#### Plotting the Rasters

We'll start making a plot of the predicted whale sightings

```{r}
ggplot() +
  geom_raster(data = cetacean_longer_df, aes(x = x, y = y, fill = richness)) +
  geom_sf(data = nat_earth_layer) +
  coord_sf(xlim = c(-125, -115), ylim = c(32,38)) +
  scale_fill_gradient(high = "red", low = "green") +
  labs(x = "Expected Whale Sighting per Unit Area",
       y = "",
       fill = "Species Richness \n (Cetacean Species per Area)",
       title = "Expected Cetacean Sitings off California's Coast") +
  theme_minimal()
```
*Figure 1: Above is a map of expected species richness of 35 cetacean species (whales and dolphins) across California's coast. We can see high richness (in red) in the Santa Barbara bight.*


